---
title: "S4 Diff Expression"
author: "Jessica, updated by Rebecca"
date: "`r Sys.Date()`"
output: html_document
---

DGE DESeq2 analysis of S4Sm using new reference genome when in competition with RE22

# Setup

# Load packages

```{r packages, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DESeq2)
library(data.table)
library(dplyr)
library(tidyr)
library(reshape2)
library(devtools)
library(ggplot2)
library(ggrepel)
library(ggfortify)
library(cluster)
library(data.table)

```

## Load data

load transcript count matrix and metadatata 
phenodata.csv file contains metadata on the count table's samples that you create separately
Make sure PHENODATA is in the same order or these commands will change data to be wrong!!!!###

```{r loaddata}

#Load Phenodata
S4_TranColData <- read.csv("phenodata.csv", header=TRUE, sep=",") %>% 
   # add column with technical replicate number based on SampleID column
   separate_wider_delim(sampleID, names=c("BioRep", "TechRep"), delim="_", cols_remove = FALSE)
head(S4_TranColData)

#Load transcript count data
S4_TranscriptCountData <- as.data.frame(read.csv("S4_transcript_count_matrix.csv", row.names="transcript_id"))
head(S4_TranscriptCountData)

```

## Clean up data

```{r dataclean}

# make sure the metadata matches the column names of the transcript count data
identical(S4_TranColData$sampleID, colnames(S4_TranscriptCountData)) # should return RTUE

# set rownames
rownames(S4_TranColData) <- S4_TranColData$sampleID

# Check all sample IDs in S4_TranColData are also in S4_TranscriptCountData and match their orders
all(rownames(S4_TranColData) %in% colnames(S4_TranscriptCountData))  # should return TRUE
all(rownames(S4_TranColData) == colnames(S4_TranscriptCountData))    # should return TRUE

```

# DeSeq2

## Make object 

```{r deseq2make}

#Set up the data matrix for S4 sample analysis in DESeq
ddsS4 <- DESeqDataSetFromMatrix(countData = S4_TranscriptCountData, 
                                  colData = S4_TranColData, 
                                  design =  ~ condition)
colData(ddsS4)
colnames(ddsS4)

```

## Collapse tech replicates

```{r deseqtechrep}

ddsS4collapse <- collapseReplicates(ddsS4, # deseq object
                                    groupby=S4_TranColData$BioRep, # grouping variable
                                    run=S4_TranColData$TechRep, # detail of runs per BioRep
                                    renameCols = TRUE)

```

## Edit deseq object

```{r editdeseq}

# Reorder the levels in ddsS4collapse so they are in the proper order for analysis
ddsS4collapse$condition <- factor(ddsS4collapse$condition, levels = c("control","treatment"))
ddsS4collapse$location <- factor(ddsS4collapse$location, levels = c("biofilm","supernatant"))

colData(ddsS4collapse)
colnames(ddsS4collapse)

```


# Session info

```{r session}

sessionInfo()

```
